---
title: "Chapter 8 - Exercise 9"
author: "Zhoumengdi Wang (zxw534)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(ISLR)
library(tree)
```

#### (a)
```{r}
set.seed(6)
training.index = sample(dim(OJ)[1], 800)

OJ_train = OJ[training.index,]
OJ_test = OJ[-training.index,]
```


#### (b)

```{r}
tree_Purchase = tree(Purchase ~ ., OJ_train)
summary(tree_Purchase)
```

This tree has 8 terminal nodes and the misclassification error rate is 14.75%.

#### (c)

```{r}
tree_Purchase
```

'*' means this is a terminal node. 
Let's focus on node number 24, the splitting value of this node is under 0.06560, which classifies a purchase as being Minute Maid (MM).
There are 60 observations in this branch, with a deviance of 77.69.
Whereas 65% of observations in this branch are indeed Minute Maid, 35% of them are not.

#### (d)
```{r}
plot(tree_Purchase)
text(tree_Purchase)
```


Most terminal nodes in the left branch are Purchase = CH.
Most terminal nodes in the right branch are Purchase = MM.
LoyalCH is involved more than others in this tree.


#### (e)
```{r}
OJ_test_pred = predict(tree_Purchase, OJ_test, type="class")
table(OJ_test$Purchase, OJ_test_pred)
```

The test error rate is (29 + 23)/(134 + 23 + 29 + 84) = 19.26%.


#### (f)
```{r}
cv_OJ = cv.tree(tree_Purchase, FUN = prune.misclass)
cv_OJ
```

#### (g)
```{r}
plot(cv_OJ$size, cv_OJ$dev, type="b")
```

#### (h)
According to these results, the tree with 5 terminal nodes has the lowest classification error rate, with 149 classification errors.

#### (i)
```{r}
pruned_OJ = prune.misclass(tree_Purchase, best=2)
plot(pruned_OJ)
text(pruned_OJ, pretty = 0)
```

#### (j)
```{r}
summary(pruned_OJ)
```

The training error rate of unpruned tree is 14.75%, the training error rate of pruned tree is 18.62%, which is higher than unpruned tree.

#### (k)
```{r}
pruned_test_pred = predict(pruned_OJ, OJ_test, type="class")
table(OJ_test$Purchase, pruned_test_pred)
```

The test error rate is (29 + 26)/(131 + 26 + 29 + 84) = 20.37%.

The teat error rate of unpruned tree is 19.26%, the test error rate of pruned tree is 20.37%, which is higher than unpruned tree.
It seems that the unpruned tree, which has more complexity structure, performance little better. 
















